---
title: "p8105_hw6_yh3964"
author: "Yirou Hu"
date: "2025-11-22"
output: github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(modelr)
library(readr)
library(p8105.datasets)
library(knitr)
library(broom)

theme_set(theme_minimal() + theme(legend.position = "bottom"))  
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```
# Problem 1

## Section a
The Washington Post has gathered data on homicides in 50 large U.S. cities and made the data available through a GitHub repository here. You can read their accompanying article here.

-   Create a `city_state` variable (e.g. "Baltimore, MD"), and a binary variable indicating whether the homicide is solved.
-   Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO -- these don't report victim race.
-   Also omit Tulsa, AL -- this is a data entry mistake.
-   For this problem, limit your analysis those for whom victim_race is white or black.
-   Be sure that victim_age is numeric.

```{r}
homicide_data = read_csv("homicide-data.csv") |>
  janitor::clean_names() |>
  mutate(city_state = paste(city, state, sep = ", "),
         resolved = as.numeric(disposition == "Closed by arrest"),
          victim_age = as.numeric(victim_age)) |>
  filter(!city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
         victim_race %in% c("White", "Black")) |>
  drop_na(victim_age)
```

## Section b
For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors.

-   Save the output of glm as an R object;
-   apply the broom::tidy to this object;
-   and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.


```{r}
baltimore_df = homicide_data |> 
  filter(city_state == "Baltimore, MD")

baltimore_glm = glm(resolved ~ victim_age + victim_sex + victim_race, 
                     data = baltimore_df, 
                     family = binomial())

baltimore_model_results = baltimore_glm |> 
  broom::tidy(conf.int = TRUE) |> 
  filter(term == "victim_sexMale") |> 
  mutate(
    OR = exp(estimate),
    lower_ci = exp(conf.low),
    upper_ci = exp(conf.high)
  )

male_female_or = 
  baltimore_model_results |> 
  pull(OR)

male_female_ci_lower =
  baltimore_model_results |> 
  pull(lower_ci) 

male_female_ci_upper = 
  baltimore_model_results |> 
  pull(upper_ci)
```

### Comments 

The estimate of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed is `r male_female_or`. These results indicate that the odds of having a solved homicide case for male is approximate 0.43 times the odds for a female. Furthermore, we are 95% confident that the true estimate is between (`r baltimore_model_results$lower_ci`, `r baltimore_model_results$upper_ci`).

## Section c

Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims.

Do this within a "tidy" pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

```{r}
model_glm <- function(df) {
  glm_model = glm(resolved ~ victim_age + victim_sex + victim_race, data = df, family = binomial() )
}

city_results = homicide_data |> 
  nest(data = -city_state) |> 
  mutate(
    models = map(data, model_glm),
    results = map(models, ~broom::tidy(.x, conf.int = TRUE, exponentiate = TRUE))
  ) |> 
  unnest(results) |> 
  filter(term == "victim_sexMale") |> 
  mutate(
    OR = exp(estimate),
    lower_ci = exp(conf.low),
    upper_ci = exp(conf.high),
    city_state = fct_reorder(city_state, OR)
  )

city_results
```


Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.

```{r}
city_results_plot =
  city_results |> 
  ggplot(aes(x = OR, y = city_state)) +
  geom_point(color = "darkblue", size = 1.5) +
  geom_errorbar(aes(xmin = lower_ci, xmax = upper_ci), width = 0.25) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "red") +
  labs(
    title = "Adjusted Odds Ratios with CI for Solving Homicides (Male vs Female Victims)",
    x = "Odds Ratio",
    y = "City"
  ) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 8))

city_results_plot
```

### Comments

-   For most cities, the ORs are less than 1, suggesting that, there are higher odds that a homicide with female will be solved compared to male after keeping all other variables fixed.

-   The cities at the top of the plot, such as Fresno, CA, and Minneapolis, MN, have ORs greater than 1, indicating a relatively higher likelihood of solving cases with male victims in these cities.

-   For cities like Tulsa, OK, and Nashville, TN, have confidence intervals that contain the OR of 1, suggesting that there is no statistically significant difference in the likelihood of solving homicides between male and female victims in those cities.

-   The confidence intervals for cities like Fresno, CA, do not cross 1, suggesting a statistically significant difference in the likelihood of solving homicides with male victims compared to female victims.

# Problem 2

For this problem, we'll use the Central Park weather data similar to data we've seen elsewhere. The code chunk below (adapted from the course website) will download these data.

```{r}
data("weather_df")
```

The boostrap is helpful when you'd like to perform inference for a parameter/ value/ summary that doesn't have an easy-to-write-down distribution in the usual repeated sampling framework. We'll focus on a simple linear regression with `tmax` as the response with `tmin` and `prcp` as the predictors, and are interested in the distribution of two quantities estimated from these data:

-   $\hat{r}^2$
-   $\hat{\beta}_1 / \hat{\beta}_2$

Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities.

-   Plot the distribution of your estimates, and describe these in words.
-   Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for $\hat{r}^2$ and $\hat{\beta}_1 / \hat{\beta}_2$.

Note: broom::glance() is helpful for extracting $\hat{r}^2$ from a fitted regression, and `broom::tidy()` (with some additional wrangling) should help in computing $\hat{\beta}_1 / \hat{\beta}_2$.

## Plots and results
```{r}
weather_clean_df = 
  weather_df |>
  select(tmax, tmin, prcp) |>
  drop_na()

bootstrap_results = 
  weather_clean_df |> 
  modelr::bootstrap(n = 5000) |> 
  mutate(
    strap = map(strap, as_tibble),
    models = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df)),
    results = map(models, \(mod) {
      r_sq = glance(mod) |> pull(r.squared)
      coef_df = tidy(mod)
      b_tmin = coef_df |> filter(term == "tmin") |> pull(estimate)
      b_prcp = coef_df |> filter(term == "prcp") |> pull(estimate)
      
      tibble(
        r_squared = r_sq,
        beta_ratio = b_tmin / b_prcp
      )
    })
  ) |> 
  select(.id, results) |> 
  unnest(results)

# Plot for R^2
p1 = bootstrap_results |> 
  ggplot(aes(x = r_squared)) +
  geom_histogram(aes(y = after_stat(density)), bins = 40, fill = "lightblue", color = "white", alpha = 0.7) +
  geom_density(color = "darkblue", size = 1) +
  labs(
    title = "Bootstrap Distribution of R-squared",
    x = "Estimated R-squared",
    y = "Density"
  ) +
  theme_minimal()

# Plot for Beta Ratio
p2 = bootstrap_results |> 
  ggplot(aes(x = beta_ratio)) +
  geom_histogram(aes(y = after_stat(density)), bins = 50, fill = "lightgreen", color = "white", alpha = 0.7) +
  geom_density(color = "darkgreen", size = 1) +
  labs(
    title = "Bootstrap Distribution of Beta Ratio (tmin / prcp)",
    x = "Estimated Ratio",
    y = "Density"
  ) +
  theme_minimal()

print(p1)
print(p2)
```

## Derive quantiles

```{r}
ci_results = 
  bootstrap_results |> 
  pivot_longer(
    cols = c(r_squared, beta_ratio),
    names_to = "term",
    values_to = "value"
  ) |> 
  group_by(term) |> 
  summarize(
    ci_lower = quantile(value, 0.025),
    ci_upper = quantile(value, 0.975)
  )

knitr::kable(ci_results, digits = 3)
```

### Comments

- We are 95% confident that the R^2 is somewhere between `r paste0("[", round(ci_results |> filter(term == "r_squared") |> pull(ci_lower), 3), ", ", round(ci_results |> filter(term == "r_squared") |> pull(ci_upper), 3), "]")` and for β₁/β₂ is `r paste0("[", round(ci_results |> filter(term == "beta_ratio") |> pull(ci_lower), 3), ", ", round(ci_results |> filter(term == "beta_ratio") |> pull(ci_upper), 3), "]")`).

- The distribution of $R²$ shows relatively stable and approximately symmetric, concentrated in the 0.935-0.947 range. This indicates the model fit ($R^2$) is highly reliable and exhibits low uncertainty across bootstrap samples.

- The distribution of $\beta_1 / \beta_2$ is strongly skewed with wide spreads. The main body of the distribution is concentrated between approximately $-100$ and $-250$, but the distribution has a long, heavy left tail extending beyond $-400$. This wide spread and pronounced skew reflect substantial instability in the ratio estimate. This instability is likely due to the denominator coefficient $\beta_{prcp}$ being small and close to zero in many bootstrap samples, which makes the ratio highly sensitive to minor variations, leading to large negative outliers.


# Problem 3

In this problem, you will analyze data gathered to understand the effects of several variables on a child’s birthweight. This dataset, available here, consists of roughly 4000 children and includes the following variables:

-   `babysex`: baby's sex (male = 1, female = 2)
-   `bhead`: baby's head circumference at birth (centimeters)
-   `blength`: baby's length at birth (centimeteres)
-   `bwt`: baby's birth weight (grams)
-   `delwt`: mother's weight at delivery (pounds)
-   `fincome`: family monthly income (in hundreds, rounded)
-   `frace`: father's race (1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other, 9 = Unknown)
-   `gaweeks`: gestational age in weeks
-   `malform`: presence of malformations that could affect weight (0 = absent, 1 = present)
-   `menarche`: mother's age at menarche (years)
-   `mheigth`: mother's height (inches)
-   `momage`: mother's age at delivery (years)
-   `mrace`: mother's race (1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other)
-   `parity`: number of live births prior to this pregnancy
-   `pnumlbw`: previous number of low birth weight babies
-   `pnumgsa`: number of prior small for gestational age babies
-   `ppbmi`: mother's pre-pregnancy BMI
-   `ppwt`: mother's pre-pregnancy weight (pounds)
-   `smoken`: average number of cigarettes smoked per day during pregnancy
-   `wtgain`: mother's weight gain during pregnancy (pounds)

## Section a

Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).

```{r}
birth_df =  
  read.csv('birthweight.csv') |>
  mutate(
    baby_sex = factor(babysex, levels = 1:2, labels = c("Male", "Female")),
    malform_present = factor(malform, levels = 0:1, labels = c("Absent", "Present")),
    mother_race = factor(mrace, 
                         levels = c(1, 2, 3, 4, 8), 
                         labels = c("White", "Black", "Asian", "Puerto Rican", "Other")),
    father_race = factor(frace, 
                         levels = c(1, 2, 3, 4, 8, 9), 
                         labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown"))
  ) 
```

Check for the missing values

```{r}
missing_tibble =
  birth_df |> 
  summarise(across(everything(), ~ sum(is.na(.x)))) |> 
  pivot_longer(everything(),
               names_to = "variables",
               values_to = "number of missing values") 

print(missing_tibble)
```

From the tibble, we know that there are no missing values in birthweight dataset.


Propose a regression model for birthweight.

-   This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two.
-   Describe your modeling process and show a plot of model residuals against fitted values -- use `add_predictions` and `add_residuals` in making this plot.

```{r}
# Proposed model
model_proposed = lm(
  bwt ~ gaweeks + blength + babysex + mheight + ppbmi + wtgain + smoken + momage + mrace, 
  data = birth_df
)

summary(model_proposed)
```

Then let's make plots about Residuals vs. Fitted Values
```{r}
resid_plot =
  birth_df |>
  add_predictions(model_proposed) |>
  add_residuals(model_proposed) |>
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.4) +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  labs(
    title = "Residuals vs. Fitted Values (Proposed Model)",
    x = "Fitted Birthweight (grams)",
    y = "Residuals (grams)"
  ) +
  theme_minimal()

print(resid_plot)
```

### Comments

- The model shows good performance with an Adjusted $R^2$ of 0.6077, explaining approximately 60.8% of the variability in birthweight. 

- The model as a whole is highly significant ($p < 2.2 \times 10^{-16}$). Most predictors are statistically significant ($p < 0.001$), with blength (birth length) showing the strongest effect ($t=59.46$). Factors like gaweeks (gestational weeks), wtgain (weight gain), and ppbmi (pre-pregnancy BMI) are positively correlated with birthweight, while smoken (smoking) is significantly associated with reduced birthweight. babysex (infant sex) is the only variable approaching significance, falling just outside the $\alpha=0.05$ level ($p=0.0897$).

- From the plots, we can observe the points are roughly evenly distributed around zero and randomly scattered around the zero line (red dashed line), supporting the assumption that the model's mean structure is reasonable. However, the plot exhibits Heteroscedasticity: the vertical spread of the residuals is much smaller for lower fitted values.

## Section b
Compare your model to two others:

- One using length at birth and gestational age as predictors (main effects only)

- One using head circumference, length, sex, and all interactions (including the three-way interaction) between these

Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.

Note that although we expect your model to be reasonable, model building itself is not a main idea of the course and we don’t necessarily expect your model to be “optimal”.

```{r}
cv_base_df = birth_df |>
  drop_na(bwt, gaweeks, blength, babysex, mheight, ppbmi, wtgain, smoken, momage, mrace, bhead)

cv_results =
  cv_base_df |>
  crossv_mc(n = 100) |> 
  mutate(
    train = map(train, as_tibble),
    test  = map(test,  as_tibble),
    
    fit_M1 = map(train, ~ lm(bwt ~ gaweeks + blength + babysex + mheight + ppbmi + wtgain + smoken + momage + mrace, data = .x)),
    fit_M2 = map(train, ~ lm(bwt ~ blength + gaweeks, data = .x)),
    fit_M3 = map(train, ~ lm(bwt ~ (bhead + blength + babysex)^3, data = .x)),
    
    rmse_M1 = map2_dbl(fit_M1, test, ~ rmse(model = .x, data = .y)),
    rmse_M2 = map2_dbl(fit_M2, test, ~ rmse(model = .x, data = .y)),
    rmse_M3 = map2_dbl(fit_M3, test, ~ rmse(model = .x, data = .y))
  )

# Summarize Mean CV-RMSE
rmse_summary =
  cv_results |>
  select(starts_with("rmse")) |>
  pivot_longer(everything(), names_to = "model", values_to = "rmse") |>
  group_by(model) |>
  summarise(
    mean_cv_rmse = mean(rmse),
    sd_cv_rmse = sd(rmse)
  )

# Display the summary table
print(rmse_summary)
rmse_summary |> knitr::kable(digits = 2)

# Plot RMSE Distribution
rmse_plot =
  cv_results |>
  select(starts_with("rmse")) |>
  pivot_longer(everything(), names_to = "model", values_to = "rmse") |>
  mutate(
    model = recode(
      model,
      "rmse_M1" = "Model 1: Proposed Model",
      "rmse_M2" = "Model 2: Length + GA",
      "rmse_M3" = "Model 3: Complex Interaction"
    ) |> 
      fct_inorder()
  ) |>
  ggplot(aes(x = model, y = rmse)) +
  geom_boxplot(aes(fill = model), alpha = 0.7, show.legend = FALSE) +
  labs(
    title = "Cross-Validated RMSE Comparison",
    x = "Model",
    y = "Root Mean Square Error (grams)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, hjust = 1))

print(rmse_plot)
```

### Comments

- The boxplot visually compares the out-of-sample prediction error (RMSE) for the three models.

- Model 1 (Proposed Model) performs moderately well, with its median RMSE higher than M3 but lower than M2, which is equal to `r round(rmse_summary[rmse_summary$model == "rmse_M1", "mean_cv_rmse"], 1)`.

- Model 2 (Length + GA) has the highest median RMSE and the widest spread of errors, indicating it is the least accurate and most variable in its predictions. The value is equal to `r round(rmse_summary[rmse_summary$model == "rmse_M2", "mean_cv_rmse"], 1)`.

- Model 3 (Complex Interaction) exhibits the lowest median RMSE and generally the lowest distribution of errors, suggesting it has the best predictive performance. The value is equal to `r round(rmse_summary[rmse_summary$model == "rmse_M3", "mean_cv_rmse"], 1)`.

- This suggests that the complex interactions in M3 significantly improve predictive accuracy compared to both the simpler M2 and the proposed M1.