---
title: "p8105_hw6_yh3964"
author: "Yirou Hu"
date: "2025-11-22"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(modelr)
library(readr)
library(p8105.datasets)
library(knitr)

theme_set(theme_minimal() + theme(legend.position = "bottom"))  
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```
# Problem 1

## Section a
The Washington Post has gathered data on homicides in 50 large U.S. cities and made the data available through a GitHub repository here. You can read their accompanying article here.

-   Create a `city_state` variable (e.g. "Baltimore, MD"), and a binary variable indicating whether the homicide is solved.
-   Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO -- these don't report victim race.
-   Also omit Tulsa, AL -- this is a data entry mistake.
-   For this problem, limit your analysis those for whom victim_race is white or black.
-   Be sure that victim_age is numeric.

```{r}
homicide_data = read_csv("homicide-data.csv") |>
  janitor::clean_names() |>
  mutate(city_state = paste(city, state, sep = ", "),
         resolved = as.numeric(disposition == "Closed by arrest"),
          victim_age = as.numeric(victim_age)) |>
  filter(!city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
         victim_race %in% c("White", "Black")) |>
  drop_na(victim_age)
```

## Section b
For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors.

-   Save the output of glm as an R object;
-   apply the broom::tidy to this object;
-   and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.


```{r}
baltimore_df = homicide_data |> 
  filter(city_state == "Baltimore, MD")

baltimore_glm = glm(resolved ~ victim_age + victim_sex + victim_race, 
                     data = baltimore_df, 
                     family = binomial())

baltimore_model_results = baltimore_glm |> 
  broom::tidy(conf.int = TRUE) |> 
  filter(term == "victim_sexMale") |> 
  mutate(
    OR = exp(estimate),
    lower_ci = exp(conf.low),
    upper_ci = exp(conf.high)
  )

male_female_or = 
  baltimore_model_results |> 
  pull(OR)

male_female_ci_lower =
  baltimore_model_results |> 
  pull(lower_ci) 

male_female_ci_upper = 
  baltimore_model_results |> 
  pull(upper_ci)
```

### Comments 

The estimate of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed is `r male_female_or`. These results indicate that the odds of having a solved homicide case for male is approximate 0.42 times the odds for a female. Furthermore, we are 95% confident that the true estimate is between (`r baltimore_model_results$lower_ci`, `r baltimore_model_results$upper_ci`).

## Section c

Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims.

Do this within a "tidy" pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

```{r}
model_glm <- function(df) {
  glm_model = glm(resolved ~ victim_age + victim_sex + victim_race, data = df, family = binomial() )
}

city_results = homicide_data |> 
  nest(data = -city_state) |> 
  mutate(
    models = map(data, model_glm),
    results = map(models, ~broom::tidy(.x, conf.int = TRUE, exponentiate = TRUE))
  ) |> 
  unnest(results) |> 
  filter(term == "victim_sexMale") |> 
  mutate(
    OR = exp(estimate),
    lower_ci = exp(conf.low),
    upper_ci = exp(conf.high),
    city_state = fct_reorder(city_state, OR)
  )
```


Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.

```{r}
city_results |> 
  ggplot(aes(x = OR, y = city_state)) +
  geom_point(color = "darkblue", size = 1.5) +
  geom_errorbar(aes(xmin = lower_ci, xmax = upper_ci), width = 0.25) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "red") +
  labs(
    title = "Adjusted Odds Ratios with CI for Solving Homicides (Male vs Female Victims)",
    x = "Odds Ratio",
    y = "City"
  ) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 8))
```

### Comments

-   For most cities, the ORs are less than 1, suggesting that, there are higher odds that a homicide with female will be solved compared to male after keeping all other variables fixed.

-   The cities at the top of the plot, such as Fresno, CA, and Minneapolis, MN, have ORs greater than 1, indicating a relatively higher likelihood of solving cases with male victims in these cities.

-   For cities like Tulsa, OK, and Nashville, TN, have confidence intervals that contain the OR of 1, suggesting that there is no statistically significant difference in the likelihood of solving homicides between male and female victims in those cities.

-   The confidence intervals for cities like Fresno, CA, do not cross 1, suggesting a statistically significant difference in the likelihood of solving homicides with male victims compared to female victims.

